### 2024-05-05 dissertation/content/p1_intro.tex
added dissertation

\chapter{The Relevance of Time-Frequency Analyses}