### 2024-05-05 dissertation/content/p2_intro.tex

\chapter{Feature Extraction with Wavelet Transforms}