### 2024-05-05 dissertation/content/p3_intro.tex

\chapter{Wavelet Scattering in Higher Dimensions}