### 2024-05-05 dissertation/main.tex

\documentclass[11pt,oneside,openany,a4paper, %... Global layout
afrikaans,UKenglish, %... Global language options
PhD, a5block %... stb-thesis options
]{stb-thesis}

\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[onehalfspacing]{setspace}
\usepackage{color}
\usepackage{parskip}%......... Paragraph spacing + zero indent
\usepackage{amssymb} % only if necessary
\usepackage{bm} %.. Bold math characters (after fonts)
\usepackage{hyperref}

\usepackage[numbers]{stb-bib}%........................... Bibliography
%.......... Auhor-year style
\addto{\captionsafrikaans}{\renewcommand{\bibname}{Lys van Verwysings}}
\addto{\captionsUKenglish}{\renewcommand{\bibname}{List of References}}

% \iftrue
% \usepackage[backref=page]{hyperref}%.... Hyperlinks & backreferences
% \else
% \usepackage[pageref]{backref}%........... Citation backreferences
% \fi
% \renewcommand*{\backrefalt}[4]{%s
% \ifcase #1 (Not cited.)%
% \or (Cited on page~#2.)%
% \else (Cited on pages~#2.)%
% \fi}

\newcommand\articleinfo[5]{%
\noindent\rule{\textwidth}{0.5pt}
\begin{center}\small\textit{#1}\par
    \begin{spacing}{0.1}
    \small Accepted in #2 on #3\par
    \small doi: #4\noindent
\end{spacing}\noindent
\end{center}\noindent
\noindent\rule{\textwidth}{0.25pt}%
\section*{\centering \textsc{Abstract}}
\begin{spacing}{1.0}    
    \small  #5 \noindent
\end{spacing}
\noindent\rule{\textwidth}{0.5pt}\pagebreak
}

\usepackage{paper1}
\usepackage{paper2}
\usepackage{paper3}

% Subappendices
\usepackage{appendix}
% Start of subappendices environment
\AtBeginEnvironment{subappendices}{%
\section*{Appendices}
\addcontentsline{toc}{section}{Appendices}
\counterwithin{figure}{section}
\counterwithin{table}{section}
}

% End of subappendices environment
\AtEndEnvironment{subappendices}{%
\counterwithout{figure}{section}
\counterwithout{table}{section}
}

\graphicspath{{./content/paper2/fig/}{./content/paper1/fig/}}

\usepackage{iftex}
\ifxetex
\usepackage[math-style=TeX, bold-style=TeX]{unicode-math}
\setmainfont{Cambria}%........................ (Windows)
\setsansfont[Scale=MatchLowercase]{Calibri}
\setmonofont[Scale=MatchLowercase]{Consolas}
\setmathfont{Cambria Math}
\defaultfontfeatures{Ligatures=TeX}
\let\bm\symbfit%.............................. Bold math
\else
\usepackage{textcomp}%........................ Additional text symbols
\usepackage[T1]{fontenc}%..................... Type 1 outline fonts
\usepackage{bm}%.............................. Bold math fonts
\fi

\begin{document}
\title{\textbf{Wavelet Scattering Tranforms Applied to Whale Vocalisations}}
\author{M.W. Rademan}{Marco Wiehann Rademan}
\faculty{Electrical and Electronic Engineering}
\degree{PhD}{Doctor of Philosophy}
\supervisor{D.J.J. Versfeld}
\cosupervisor{J.A. Du Preez}
% \date{August}{2024} 
\SetCopyrightOff
\address{\AorE{Departement Siviele Ingenieurswese,\\ Stellenbosch Universiteit,\\
Privaatsak X1, 7602 Matieland, Suid Afrika.
}{%
Department of Civil Engineering,\\ Stellenbosch University,\\
Private Bag X1, 7602 Matieland, South Africa.}}

% \frontmatter
\begin{singlespace}
\TitlePage%..................... Single spaced title page
\end{singlespace}

\begin{singlespace}
\DeclarationDate{\today}
\DeclarationPage
\end{singlespace}
% \DeclarationPage

\begin{abstract}

\end{abstract}

\begin{abstract}[afrikaans]

\end{abstract}

\tableofcontents
\listoffigures
\listoftables
\chapter*{Nomenclature}
\label{chap:nomenclature}
\addcontentsline{toc}{chapter}{\nameref{chap:nomenclature}}

% \begin{singlespace}
\input{content/introduction}

\input{content/p1_intro}
\input{content/paper1/main}

\input{content/p2_intro}
\input{content/paper2/main}

\input{content/p3_intro}
\input{content/paper3/main}
% \end{singlespace}

\bibliographystyle{stb-bib-eng-a}
\bibliography{references}

\end{document}