### 2024-05-06 dissertation/content/p1_intro.tex

\chapter{The Applications of Time-Frequency Analyses}