### 2024-05-14 dissertation/content/nomenclature.tex
started intro to p3

\section*{Notation}
Some notations are used interchangably, which is obvious from the context. Due to the nature of the publication format, notation is explained within the article it occurs.

\begin{tabular}{p{0.15\textwidth} p{0.75\textwidth}}
    $f(x)$ & a continuous function of $x$ \\
    $f(x, y)$ & a continuous multi-variable function of $x$ and $y$ \\
    $f[n]$ & a discrete-time (sampled) function indexed by $n$ \\
    $f[n, m]$ & a multi-variable discrete-time function indexed by $n$ and $m$ \\
    $\mathcal{F} \{f\}$ & the Fourier transform operator acting on $f$\\
    $\hat{f}$ & the Fourier transform of $f$ or an estimate of $f$, obvious from context\\
    $x^*$ & the complex conjugate of $x$\\
    $p(x)$ & a probability density \\
    $p(x|y)$ & a conditional probability density \\
    $\{a, b, ...\}$ & a set of numbers\\
    $A \setminus B$ & set minus: the set $A$ excluding the elements of $B$\\
    $A \times B$ & Cartesian product of sets\\
    $| x |$ & magnitude of a number (modulus), or size of a set\\
    $\mathbb{R}$ & the set of real numbers\\
    $\mathbb{C}$ & the set of complex numbers\\
    $\mathbb{N^+}$ & the set of natural numbers (excluding 0)\\
    $\mathbb{Z}$ & the set of integers\\
    $x \sim p$ & $x$ is drawn from the probability density $p$\\
    $\frac{d}{dx} f$ & derivative with respect to $x$ of a function $f$\\
    $\mathcal{U}_j x$ & scalogram of the $j$'th level of $x$\\
    $\mathcal{S}_j x$ & scattering coefficients of the $j$'th level of $x$\\
    $x * y$ & convolution of two functions\\
    $x \otimes y$ & circular convolution of two discrete functions\\
    $(x)_{\downarrow r}$ & a discrete function subsampled by a factor $r$\\
    $\mathbf{x}$ & a vector\\
\end{tabular}

\begin{tabular}{p{0.15\textwidth} p{0.75\textwidth}}
    $X[k,m]$ & a time-frequency decomposition of $x$, with $k$ the frequency index and $m$ the time index \\
    $S[k,m]$ & power spectrum calculated from $X[k, m]$ \\
    $H[m]$ & spectral entropy calculated from $S[k, m]$ \\
    $X(s, b)$ & continuous wavelet transform of $x$, with $s$ as scale and $b$ as translation \\
    $\uppsi_\lambda$ & sometimes used as short-hand for  $\uppsi(\lambda t)$\\
    $\text{RM}^F_m (x)$ & rolling median filter operator with a window of size $F$ operating on index variable $m$ on a discrete function $x$\\
    $\text{RA}^F_m (x)$ & rolling average filter operator with a window of size $F$ operating on index variable $m$ on a discrete function $x$\\
    $x \triangleq y$ & $x$ is defined as $y$\\
\end{tabular}

\clearpage
\section*{Common Variables}
Some variables are used interchangably, which is obvious from the context. Due to the nature of the publication format, variables are explained within the article it occurs.

\begin{tabular}{p{0.15\textwidth} p{0.75\textwidth}}
    a & Hello \\
    a & Hello \\
    a & Hello \\
    a & Hello \\
\end{tabular}

\section*{Acronyms}
\begin{tabular}{p{0.15\textwidth} p{0.75\textwidth}}
    GMM & Gaussian mixture model \\
    DT & discrete-time \\
    POI & point(s) of interest \\
    EM & expectation maximization \\
    SE & spectral entropy \\
    PAM & passive acoustic monitoring \\
    PSD & power spectral density \\
    STFT & short-time Fourier transform \\
    TPR & true positive rate \\
    FPR & false positive rate \\
    PDF & probability density function \\
    SNR & signal-to-noise ratio \\
    AMFM & amplitude and frequency modulated \\
    BLED & band-limited energy detection \\
    TF & time-frequency \\
    CWT & continuous wavelet transform \\
    DWT & discrete wavelet transform \\
    SWT & stationary wavelet transform \\
    WPD & wavelet packet decomposition \\
    MODWT & maximal overlap discrete wavelet transform \\
    MFCC & Mel-frequency cepstral coefficient \\
    DNN & deep neural network \\
    CNN & convolutional neural network \\
    HMM & hidden Markov model \\
    SOTA & state-of-the-art \\
    LSTM & long-short-term memory \\
    EWT & Endangered Wildlife Trust \\
    WS & wavelet scattering \\
    WST & wavelet scattering transform \\
    FFT & fast Fourier transform \\
    GPU & graphics processing unit \\
    ROC & receiver operating characteristic \\
    CLS & cosine-log scattering \\
    LDA & linear discriminant analysis \\
    ML & machine learning \\
    NN & neural network \\
    SVM & support vector machine \\
\end{tabular}

\begin{tabular}{p{0.15\textwidth} p{0.75\textwidth}}

    1D & 1-dimensional \\
    2D & 2-dimensional \\
    SIFT & scale-invariant feature transform \\
    LPF & low-pass filter \\
    IR & impulse response \\
    FIR & finite impulse response \\
    AUC & area under curve \\
    SD & standard deviation \\
    STPR & soft true positive ratio \\
    STNR & soft true negative ratio \\
    BPF & band-pass filter \\

\end{tabular}

% TODO: paper 1 acronyms and consolodate