### 2024-05-21 dissertation/content/introduction.tex

\chapter{Introduction}

\Ac{ml} models have become mainstay in software products and automation pipelines. Often, when there is processing power available, a \ac{ml} model will be deployed to increase efficiency, accuracy and limit human interaction. Various models, whether they are statistical, Bayesian or probabilistic, decision trees or \acfp{nn}, are configured to assist or perform various tasks across many industries.

Modern research is very focussed on advancing \ac{ml} integration, improving model performance and investigating new techniques. A large branch of this research is particularly focussed on data-driven approaches using \acp{nn}. The collection of data therefore becomes an important part of model training and deployment. \Acp{nn} present some challenges and concerns, since they require vast amounts of data, and their operation is often unpredictable and not well understood \citep{nnsurvey}. They are typically treated as ``black box'' functions, which researchers improve and design through trial and error via hyper-parameter optimisation and architecture exploration.

In the field of \ac{pam}, which focuses on monitoring wildlife using non-invasive acoustic sensors, the majority of recent \ac{sota} models are \ac{nn}-based \citep{pamsurvey}. However, metrics of real-life deployment and practical model evaluation are often ignored, since the acoustical landscape of natural environments can be surprisingly varied. For example, one cannot expect a \ac{nn} model which was trained on data that does not contain rainfall noise, to effectively operate when it starts raining. This is especially relevant for \acp{cnn}, since the learnt filters are often unexpectedly shaped in the frequency domain \cite{sincnet}.

Long moments of silence, weather patterns and migrating animals contribute to an ever-changing soundscape which can cause many false positives in naively trained and deployed models \citep{pamguide}. Systems which aim to track specific species are often faced with this ``needle in a haystack'' problem, in which the sounds you are looking for are rare compared to the lifetime of the deployed sensors. As such, human intervention with recorded audio is a must-have when trying to reliably track species via audio signals, which can be a very laborious task indeed.

As such, designing models and systems which are robust to all the aforementioned factors are critical in automating wildlife conservation studies based on non-invasive audio sensors. This leads to the natural question of the suitability of \acp{nn} in this field and to what extent they can be used to be effective.

At the core of many audio-based systems lies a \ac{tf}-decomposition, typically \acp{mfcc} or the \ac{stft}. This dissertation focuses on \ac{tf} decompositions, and how they can be leveraged in order to reduce the usage or even entirely replace \acp{nn} in \ac{pam} systems. At the end of this investigative journey, we naturally arrive an entirely new generalisation of a feature extraction method based on \ac{ct} wavelets, which can be utilised as a drop-in replacement for a \ac{cnn} front-end. Wavelet scattering features be further improved with deep learning across filter channels \cite{nnphasecollapse}. The integration and investigation of wavelet-based processing in \ac{pam} systems are relatively scarce, even though it is very prevalent in other fields \citep{denoising,ws_fault_diag, waveletfault, waveletflight, waveletflow, waveletimagedenoising}, which we aim to remedy in the publications presented in this dissertation.

\section{Problem Statement}
\label{sec:problem}

In summary, some of the many challenges posed by \ac{pam} systems are:
\begin{enumerate}
    \item Long periods of silence or noise, leading to unbalanced data classes which can have a tendency for creating many false positives in models if not trained correctly.
    \item Constantly evolving noise landscape, with the interference of other bioacoustical sounds which we are often not interested in classifying.
    \item Large audio databases and recordings, which motivates the use of fast algorithms and processing techniques.
    \item Poor \ac{snr}, often less than 0 dB.
    \item Incomplete or mislabeled training data, due to poor \ac{snr} and human error.
    \item Small datasets, which only cover a portion of environment variability, making it unsuitable for a general model.
    \item Non-white, non-stationary background noise, which makes classification and signal detection non-trivial tasks.
    \item Poorly understood species behaviour in terms of their bioacoustics and movements.
\end{enumerate}

All the above factors seem demotivate the use of \acp{nn} for the following reasons: principle of operation can be unpredictable; computation time; requires reliable training data or special techniques to mitigate problematic data. Although it is possible to address these challenges, it is worth exploring non-\ac{nn} methods.

As a reponse to the challenges posed by \ac{pam} systems, we focus specifically on alternative approaches to \acp{nn}, by considering the modification of traditional classifiers and detectors with wavelet-based \ac{tf}-decompositions.

\section{Layout}

As this is a dissertation by publication, each paper can be read as a stand-alone item, with all the necessary literature and knowledge captured within. As such, this dissertation is structured with additional contextual chapters (chapters \ref*{chap:p1i}, \ref*{chap:p2i}, \ref*{chap:p3i}) prior to each publication (chapters \ref*{chap:p1}, \ref*{chap:p2}, \ref*{chap:p3}).

Each contextual chapter is in itself an introduction to the literature and concepts presented in the following chapter. Contextual chapters serve to unify the content of the published papers and show the natural progression \ac{tf} decompositions, which finally arrives at a new wavelet scattering transform. All relevant literature is therefore gradually introduced, due to each paper conceptually expanding on the previous paper's ideas and contributions.

Although this dissertation is mostly focussed on the application to \ac{pam} systems, the theory and findings it presents are general. As such, the contextual chapters are generalised mathematical descriptions of the concepts present in each paper. Papers 1 and 2 (chapters \ref*{chap:p1}, \ref*{chap:p2}) are focussed on applying it to \ac{pam} systems. The final paper (chapter \ref*{chap:p3}) shows a generalised result from all the knowledge and investigations presented in prior papers.

Each paper is presented as-is, with no modifications, save for unifying mathematical notation and the renumbering of items.

\section{Journal Papers and Contributions}

The three papers presented in this dissertation and their contribution in relation to \ac{pam} and general \ac{ml} is described below. Note that the actual journal article titles are used, which are different from the chapter titles.

\begin{enumerate}
    \item \textbf{Chapter \ref*{chap:p1}: Soft-output Signal Detection for Cetacean Vocalizations using Spectral Entropy, K-means Clustering and the Continuous Wavelet Transform} -- we consider the use of the \ac{cwt} as a \ac{tf} decomposition for \ac{se} detectors, where the \ac{stft} is typically used. We then extend the conventional detector to be of a probabilistic nature. This paper introduces wavelet-based \ac{tf}decompositions in this dissertation and lays the groundwork for further extension and investigation.
    \item \textbf{Chapter \ref*{chap:p2}: Detecting and classifying Whale Calls with Wavelet Scattering and Spectral Entropy} -- we build on the work of chapter \ref*{chap:p1}, and instead consider wavelet scattering as a \ac{tf} decompistion, for which the first-level coefficients can be considered as a ``blurred'' \ac{cwt}. We further improve the \ac{se} detector using \acp{gmm}, and integrate the detector system with a classifier using wavelet scattering features. This paper focusses on the practicality of combining and improving older methods, while also introducing wavelet scattering to the field of \ac{pam}.
    \item \textbf{Chapter \ref*{chap:p3}: Separable Wavelet Scattering} -- following from the ideas presented in chapters \ref*{chap:p1} and \ref*{chap:p2}, especially with reference to computational efficiency discussed in chapter \ref*{chap:p2}, an entirely new type of wavelet scattering is generalised. We evaluate the new wavelet scattering technique with the MNIST and MedMNIST datasets for quick evaluation of the method compared to other similar wavelet scattering decompositions. We consider this paper to be the largest contribution presented in this dissertation, due to its general nature, whose design stemmed from the needs of \ac{pam} systems as discussed in section \ref*{sec:problem}.
\end{enumerate}

I am the main author in all the papers presented in this dissertation, in which I conducted all relevant research, performed the writing and submitted to the journals. Each chapter containing a publication declares detailed author contributions in terms of the CRediT system.