### 2024-05-21 dissertation/main.tex

\documentclass[11pt,oneside,openany,a4paper, %... Global layout
afrikaans,UKenglish, %... Global language options
PhD, a5block %... stb-thesis options
]{stb-thesis}

\usepackage[english]{babel}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage[onehalfspacing]{setspace}
\usepackage{color}
\usepackage{parskip}%......... Paragraph spacing + zero indent
\usepackage{amssymb} % only if necessary
\usepackage{bm} %.. Bold math characters (after fonts)
\usepackage{hyperref}

\usepackage[numbers]{stb-bib}%........................... Bibliography
%.......... Auhor-year style
\addto{\captionsafrikaans}{\renewcommand{\bibname}{Lys van Verwysings}}
\addto{\captionsUKenglish}{\renewcommand{\bibname}{List of References}}

% \iftrue
% \usepackage[backref=page]{hyperref}%.... Hyperlinks & backreferences
% \else
% \usepackage[pageref]{backref}%........... Citation backreferences
% \fi
% \renewcommand*{\backrefalt}[4]{%s
% \ifcase #1 (Not cited.)%
% \or (Cited on page~#2.)%
% \else (Cited on pages~#2.)%
% \fi}

\newcommand\articleinfo[5]{%
\noindent\rule{\textwidth}{0.5pt}
\begin{center}\small\textit{#1}\par
    \begin{spacing}{0.1}
    \small Accepted in #2 on #3\par
    \small doi: #4\noindent
\end{spacing}\noindent
\end{center}\noindent
\noindent\rule{\textwidth}{0.25pt}%
% \pagebreak
\section*{\centering \textsc{Abstract}}
\begin{spacing}{1.0}    
    \small  #5 \noindent
\end{spacing}
\clearpage
}

\usepackage{paper1}
\usepackage{paper2}
\usepackage{paper3}

% Subappendices
\usepackage{appendix}
% Start of subappendices environment
\AtBeginEnvironment{subappendices}{%
\section*{Appendices}
\addcontentsline{toc}{section}{Appendices}
}

% End of subappendices environment
\AtEndEnvironment{subappendices}{%
}

% Chapter format
\usepackage{titlesec}
\titleformat{\chapter}
  {\LARGE\bfseries}{\thechapter}{1em}{}
\titlespacing*{\chapter}{0pt}{3.5ex plus 1ex minus .2ex}{2.3ex plus .2ex}

\graphicspath{{./content/paper2/fig/}{./content/paper1/fig/}}

\usepackage{iftex}
\ifxetex
\usepackage[math-style=TeX, bold-style=TeX]{unicode-math}
\setmainfont{Cambria}%........................ (Windows)
\setsansfont[Scale=MatchLowercase]{Calibri}
\setmonofont[Scale=MatchLowercase]{Consolas}
\setmathfont{Cambria Math}
\defaultfontfeatures{Ligatures=TeX}
\let\bm\symbfit%.............................. Bold math
\else
\usepackage{textcomp}%........................ Additional text symbols
\usepackage[T1]{fontenc}%..................... Type 1 outline fonts
\usepackage{bm}%.............................. Bold math fonts
\fi

\begin{document}
\title{\AorE{Wavelet Scattering Transformasies Toegepas op Walvis Vokaliserings}{Wavelet Scattering Transforms Applied to Whale Vocalisations}}
\author{M.W. Rademan}{Marco Wiehann Rademan}
\faculty{Electrical and Electronic Engineering}
\degree{PhD}{Doctor of Philosophy}
\supervisor{D.J.J. Versfeld}
\cosupervisor{J.A. Du Preez}
% \date{August}{2024} 
\SetCopyrightOff
\address{\AorE{Departement Elektriese en Elektroniese Ingenieurswese,\\ Stellenbosch Universiteit,\\
Privaatsak X1, 7602 Matieland, Suid Afrika.
}{%
Department of Electrical and Electronic Engineering,\\ Stellenbosch University,\\
Private Bag X1, 7602 Matieland, South Africa.}}

% \frontmatter
\begin{singlespace}
\TitlePage%..................... Single spaced title page
\end{singlespace}

\begin{singlespace}
\DeclarationDate{\today}
\DeclarationPage
\end{singlespace}
% \DeclarationPage

\begin{abstract}

In this dissertation by publication, we investigate the application of wavelet scattering transforms on whale vocalisations, by following the natural progression of 3 consecutive papers.

We consider how wavelets can be used to replace the short-time Fourier transform (STFT) in signal detection. Wavelet scattering is introduced to the field of passive acoustic monitoring as a convolutional neural network (CNN) alternative for small datasets, which is combined with an improved entropy-based signal detector.

For 1-dimensional (1D) signals, wavelet scattering can be considered as a generalisation of Mel-frequency cepstral coefficients (MFCCs), and can typically outperform MFCCs in audio classification applications. In general, wavelet scattering filtering operations resemble the first layers of a CNN. An effective feature extractor may be constructed with wavelet scattering filters without requiring any learning. The main contribution of this dissertation is the introduction of separable filters for a wavelet scattering transform, which is generalised to an arbitrary number of dimensions.

\end{abstract}

\begin{abstract}[afrikaans]

  In hierdie publikasie proefskrif, ondersoek ons die toepassing van wavelet scattering transformasies op walvis vokaliserings, deur middel van opeenvolgende joernaal arikels.

  Ons orweeg hoe wavelets gebruik kan word om die kort-tyd Fourier transformasie (KTFT) in sein deteksie te vervang. Wavelet scattering is bekendgestel in die veld van passiewe akoustiese monitering, as an alternatief to konvolusionele neurale netwerke (KNNs). Verder, kombineer ons wavelet scattering met 'n klassieke entropie-gebasseerde sein detektor wat ons verbeter met addisionele tegnieke.

  Vir eendimensionele seine kan wavelet scattering as 'n veralgemening van Mel-frekwensie kepstrale koeffisi\"{e}nte (MFKKs) beskou word, altans 'n verbeterde weergawe daarvan. In die algemeen, is wavelet scattering vergelykbaar met die eerste lae van KNNs. Die hoof bydrag van hierdie proefskrif is die bekendstelling van skeibare filters vir die wavelet scattering transform, wat veralgemeen is vir 'n arbitr\^{e}re aantal intreedimensies.

\end{abstract}

\tableofcontents
\listoffigures
\listoftables
\chapter*{Nomenclature}
\label{chap:nomenclature}
\markboth{Nomenclature}{\uppercase{Nomenclature}}
\addcontentsline{toc}{chapter}{\nameref{chap:nomenclature}}
\input{content/nomenclature.tex}

% \begin{singlespace}
\input{content/introduction}

\input{content/p1_intro}
\input{content/paper1/main}

\input{content/p2_intro}
\input{content/paper2/main}

\input{content/p3_intro}
\input{content/paper3/main}

\input{content/conclusion}
% \end{singlespace}

\bibliographystyle{stb-bib-eng-a}
\bibliography{references_consolodated}

\end{document}