### 2024-05-27 dissertation/content/p2_intro.tex

\acrodef{dct}[DCT]{discrete cosine transform}

\chapter{Feature Extraction with Wavelet Transforms}
\label{chap:p2i}

\ac{tf} decompositions are often used to extract features from signals. In audio, \acp{mfcc} are features calculated from the \ac{stft} of the signal. Note that a ``\ac{tf}'' decomposition, as referred to in this dissertation, does not necesarrily require a time variable: spatial variables are often present instead. In this sense, a \ac{cnn} is can also be considered as a ``\ac{tf}'' decomposition, where multi-dimensional index variables (or index vectors) are used for ``time'' and frequency:
\begin{equation}
    X[\vect{k}, \vect{m}] = \left( x * \upsilon_{\vect{k}} \right)_{\downarrow \vect{d}}, \ \upsilon_{\vect{k}} \in \Upsilon.
\end{equation}

In this case, we abuse notation to indicate that the operator $(\cdot)_{\downarrow \vect{d}}$ downsamples by different amounts for each dimension of $\vect{m}$.

The downsampling factor $\vect{d}$ limits the bandwidth of $X$, which, as discussed in chapter \ref{chap:p1i}, is selected so as to retain all information carried by the filters in $\Upsilon$. Although not notated, different filters (different $\vect{k}$) may have different values of $\vect{d}$. Notation to support such a case is neglected for clarity, but is later expanded upon in chapter \ref{chap:p3}.

\section{The Importance of Invariance}

The decomposition $X$ is often further processed by additional filters $\Phi_i[\vect{k}, \vect{m}]$, where $i$ is the index of the post-processing filter:
\begin{equation}
    \tilde{X_i}[\vect{k}, \vect{m}] = X * \Phi_i.
\end{equation}
The filters $\Phi_i$ are not necesarrily multi-dimensional, and may only operate along a single dimension of $X$. If $\Phi_i$ is an averaging (low-pass) filter, some additional stability can be observed in $\tilde{X_i}$ compared with $X$. This allows for the introduction of invariances to various deformations applied to $X$, which can be greatly beneficial for some applications \citep{cnninvariance}. In a sense, cascased structures of $X$ and $\tilde{X_i}$ are defined in a \ac{cnn} for which the network learns the required invariances \cite{2dscattering}.

It is widely known that enforcing invariance in feature extraction and \ac{ml} pipelines increase performance. Enforcing invariance in \acp{nn} has been shown to improve the network, while also improving the interpretability of the network \cite{cnninvariance2}. For example, when classifying images, invariance to scale, shear, rotation and translation deformations, de\-pending on image properties, may be useful when extracting features. Specifically, defor\-mations of a specified order should approximately map to the same point in feature space:
\begin{equation}
    \Gamma\left(\delta\left(x\left[\vect{n}\right]\right)\right) \approx \Gamma\left(x\left[\vect{n}\right]\right),
\end{equation}
where $\Gamma$ is the feature extraction operator, and $\delta$ is the deformation operator applied to $x$.

Invariance filters ($\Phi_i$) can also have additional properties that makes it favourable for audio applications. For example, many applications ``blur'' (low-pass) the TF decomposition \citep{seyicwt, speechmodulation}, which can reduce the impact of noise fluctuations, and can also serve to suppress small frequency fluctuations across adjascent frequency bins (if the blurring kernel $\Phi_i$ filters across frequency as well). The blurring operation can improve the performance of signal detectors and classifiers. However, as a cost, time localisation and/or frequency localisation is reduced.

A new problem is introduced when enforcing invariances via the $\Phi_i$ filters - information is lost via averaging \cite{1dscattering1}. Various techniques can be used to further retain information, while also retaining invariance. Wavelet scattering is such a method \cite{ws,2dscattering}. In a sense, multilayer \acp{cnn} topologies with skip connections can be thought of as a method to recover lost information in the deeper layers.

\section{MFCCs Reframed as Shift-Invariant Features}

\acp{mfcc} are the most widely used feature extraction method for audio signals \citep{mfccreview}, save for \ac{cnn} frontends which also act as a feature extractor when operating on the \ac{stft}. For this reframing, we ignore the \ac{dct} applied to the filtered values, since this linear transformation only serves to decorrelate the filter coefficients \citep{dctdecorrelation}.

The magnitude spectrum used in \ac{mfcc} calculation is expressed as a \ac{tf} decomposition as defined by equations (\ref*{eqn:gentf}) and (\ref*{eqn:stftfilter}):
\begin{equation}
    X[l, m] = |\upsilon_l * x|_{\downarrow d},
\end{equation}
where $\upsilon_l$ is the $l$'th \ac{stft} filter.

The Mel spectrogram \ac{tf} decomposition $X_\text{Mel}$ is then calculated as
\begin{equation}
    X_\text{Mel}[k, m] = \log\left(X[l, m] * \Phi_k[l]\right) \bigg|_{l = 0},
\end{equation}
where $\Phi_k$ is the $k$'th MFFC triangular filter operating along the STFT index $l$. The evaluation at $l=0$ indicates that this operation is a multiplication and summation only.

We can view this process as a \ac{tf} decomposition with STFT filters $\upsilon_l \in \Upsilon$, which is modified with the Mel-scale triangular filters $\Phi_k$. $\upsilon_l$ provides time-shift invariance due to the STFT window, whereas $\Phi_k$ selectively averages frequency content of the STFT, thereby providing some frequency-shift invariance.

\section{Wavelets as a MFCC Generalisation}

We can construct a wavelet filterbank which introduces time and frequency-shift invariance in a similar manner to the Mel spectrogram. This process is constructed in the opposite order (frequency-shift, then time-shift invariance) to the Mel spectrogram. However, its properties remain similar \cite{ws}.

We construct wavelet band-pass filters $\upsilon_k \in \Upsilon$ which is spaced in frequency such that the desired invariance properties is obtained (as discussed in section \ref*{sec:p1i:wavelets}). We then utilise a single low-pass filter $\Phi[\vect{m}]$ which then provides invariance in the time and/or spatial dimensions:
\begin{gather}
    \label{eqn:general_scalogram}
    X[\vect{k}, \vect{m}] = \left|x * \upsilon_{\vect{k}}\right|_{\downarrow \vect{d}}, \ \upsilon_{\vect{k}} \in \Upsilon; \\
    \label{eqn:general_scattering}
    \tilde{X}[\vect{k}, \vect{m}] = X[\vect{k}, \vect{m}] * \Phi[\vect{m}]
\end{gather}

The notation in equations (\ref*{eqn:general_scalogram}) and (\ref*{eqn:general_scattering}) conforms the general \ac{tf} description of chapter \ref*{chap:p1i}, although the specific wavelet scattering operators in chapters \ref*{chap:p2} to \ref*{chap:p3} follow standard notation \citep{waveletbook}. These operators are known as wavelet scattering, which formalise the notion of ``averaging a \ac{tf}-decomposition'' in terms of signal processing theory. Chapters \ref*{chap:p2} to \ref*{chap:p3} discuss the specifics, flavours and implementation of scattering operators at length.